\bf Properties and Bayesian fitting of restricted Boltzmann machines

### 2018-09-14 writing/paper-rbm/chapters/rbm.Rmd
try to clear up the discussion

In addition to such degeneracy, we find that RBMs can easily exhibit types of instability, related to how sensitive the model probabilities can be to small changes in data outcomes. In practice, this may be seen when a single pixel change in an image results in a wildly different classification in an image classification problem. Occurrences of such behavior have recently been documented in RBMs [@li2014biclustering], as well as other deep architectures [@szegedy2013intriguing; @nguyen2014deep]. We describe (related) potential issues of model instability, degeneracy and uninterpretability for the RBM class (which are related properties) and examine the presence of these in Section \@ref(explorations-of-model-properties-through-simulation) through simulations of small, manageable examples.

@@

RBM models constitute an interesting class of undirected graphical models that are thought to be useful for supervised learning tasks. However, when viewed as generative statistical models, RBMs cn be susceptible to forms of model impropriety such as near-degeneracy, S-instability, and uninterpretability. Additionally, these models are difficult to fit using a rigorous methodology tht cn incorporate estimtion uncertainty, due to the dimension of the parameter space coupled with the size of the latent variable space.

In this paper, we have presented three fully Bayes-principled MCMC-based methods for fitting RBMs, which provide posterior distributions for quantifying uncertainty in estimating model parameters and probabilities (e.g., Figure \@ref(fig:fitting-plot)). Common practice is to use a kind of MCMC to overcome fitting complexities. However because of the size of the space to be filled with MCMC iterates, convergence and mixing of these methods will be slow. Marginalization over the latent variables in the model is shown to improve mixing, but is numerically intractable due to the necessity of repeated calculation of the normalizing constant. Due to the extreme flexibility in this model class, rigorous likelihood-based fitting for a RBM can typically seek to merely merely return the (discrete) empirical distribution for visibles, and practitioners should be aware of this and employ some form of regularization or penalization in the fitting.

>>>>

In addition to such degeneracy, we find that RBMs can easily exhibit types of instability, related to how sensitive the model probabilities can be to small difference in data outcomes. In practice, this may be seen when a single pixel change in an image results in a wildly different classification in an image classification problem. Occurrences of such behavior have recently been documented in RBMs [@li2014biclustering], as well as other deep architectures [@szegedy2013intriguing; @nguyen2014deep]. We describe (related) potential issues of model instability, degeneracy and uninterpretability for the RBM class (which are related properties) and examine the presence of these in Section \@ref(explorations-of-model-properties-through-simulation) through simulations of small, manageable examples.

@@

RBM models constitute an interesting class of undirected graphical models that are thought to be useful for supervised learning tasks. However, when viewed as generative statistical models, RBMs can be susceptible to forms of model impropriety such as near-degeneracy, S-instability, and uninterpretability. These model instability problems relate to how useful the model may be for representing realistic data generation mechanisms. In this paper, we have provided a theoretical framework for discussing the model properties of RBMs, as well as provided empirical evidence that these model problems can arise away from the origin of the parameter space.

Additionally, these models are difficult to fit using a rigorous methodology that can incorporate estimation uncertainty, due to the dimension of the parameter space coupled with the size of the latent variable space. We have presented three fully Bayes-principled MCMC-based methods for fitting RBMs, which provide posterior distributions for quantifying uncertainty in estimating model parameters and probabilities (e.g., Figure \@ref(fig:fitting-plot)) while aiming to avoid areas of the parameter space that will potentially lead to model impropriety. We have shown small examples where this fitting methodology does work, but becomes quickly intractable in large data cases.

The current common practice is to use a kind of MCMC to overcome fitting complexities, and due to the extreme flexibility in this model class, rigorous likelihood-based fitting for a RBM can typically seek to merely merely return the (discrete) empirical distribution for visibles. Current fitting methodology for RBMs does not provide uncertainty quantification for the fitted parameters, and it is unclear that they seeks to address the issues in model impropriety detailed in this paper. Practitioners should be aware of this and employ some form of regularization or penalization in the fitting.

<!-- because of the size of the space to be filled with MCMC iterates, convergence and mixing of these methods will be slow. Marginalization over the latent variables in the model is shown to improve mixing, but is numerically intractable due to the necessity of repeated calculation of the normalizing constant.  -->

### 2018-09-14 writing/paper-rbm/docs/paper.tex
try to clear up the discussion

\par
\begin{figure}
\includegraphics[width=1\linewidth]{paper_files/figure-latex/sample-models-1} \includegraphics[width=1\linewidth]{paper_files/figure-latex/sample-models-2} \caption{Ten 4-pixel images simulated from a degenerate model (a) compared to ten 4-pixel images simulated from a non-degenerate model (b). The degenerate model places almost all probability on one outcome, causing the image to be generated repeatedly, whereas the non-degenerate model shows more realistic variation.}\label{fig:sample-models}
\end{figure}
In addition to such degeneracy, we find that RBMs can easily exhibit
types of instability, related to how sensitive the model probabilities
can be to small changes in data outcomes. In practice, this may be seen
when a single pixel change in an image results in a wildly different
classification in an image classification problem. Occurrences of such
behavior have recently been documented in RBMs (Li
\protect\hyperlink{ref-li2014biclustering}{2014}), as well as other deep
architectures (Szegedy et al.
\protect\hyperlink{ref-szegedy2013intriguing}{2013}; Nguyen, Yosinski,
and Clune \protect\hyperlink{ref-nguyen2014deep}{2014}). We describe
(related) potential issues of model instability, degeneracy and
uninterpretability for the RBM class (which are related properties) and
examine the presence of these in Section
\ref{explorations-of-model-properties-through-simulation} through
simulations of small, manageable examples.

@@

RBM models constitute an interesting class of undirected graphical
models that are thought to be useful for supervised learning tasks.
However, when viewed as generative statistical models, RBMs cn be
susceptible to forms of model impropriety such as near-degeneracy,
S-instability, and uninterpretability. Additionally, these models are
difficult to fit using a rigorous methodology tht cn incorporate
estimtion uncertainty, due to the dimension of the parameter space
coupled with the size of the latent variable space.

In this paper, we have presented three fully Bayes-principled MCMC-based
methods for fitting RBMs, which provide posterior distributions for
quantifying uncertainty in estimating model parameters and probabilities
(e.g., Figure \ref{fig:fitting-plot}). Common practice is to use a kind
of MCMC to overcome fitting complexities. However because of the size of
the space to be filled with MCMC iterates, convergence and mixing of
these methods will be slow. Marginalization over the latent variables in
the model is shown to improve mixing, but is numerically intractable due
to the necessity of repeated calculation of the normalizing constant.
Due to the extreme flexibility in this model class, rigorous
likelihood-based fitting for a RBM can typically seek to merely merely
return the (discrete) empirical distribution for visibles, and
practitioners should be aware of this and employ some form of
regularization or penalization in the fitting.

>>>>

\par
\begin{figure}
\includegraphics[width=1\linewidth]{paper_files/figure-latex/sample-models-1} \includegraphics[width=1\linewidth]{paper_files/figure-latex/sample-models-2} \caption{Ten 4-pixel images simulated from a degenerate model (a) compared to ten 4-pixel images simulated from a non-degenerate model (b). The degenerate model places almost all probability on one outcome, causing the image to be generated repeatedly, whereas the non-degenerate model shows more realistic variation.}\label{fig:sample-models}
\end{figure}
In addition to such degeneracy, we find that RBMs can easily exhibit
types of instability, related to how sensitive the model probabilities
can be to small difference in data outcomes. In practice, this may be
seen when a single pixel change in an image results in a wildly
different classification in an image classification problem. Occurrences
of such behavior have recently been documented in RBMs (Li
\protect\hyperlink{ref-li2014biclustering}{2014}), as well as other deep
architectures (Szegedy et al.
\protect\hyperlink{ref-szegedy2013intriguing}{2013}; Nguyen, Yosinski,
and Clune \protect\hyperlink{ref-nguyen2014deep}{2014}). We describe
(related) potential issues of model instability, degeneracy and
uninterpretability for the RBM class (which are related properties) and
examine the presence of these in Section
\ref{explorations-of-model-properties-through-simulation} through
simulations of small, manageable examples.

@@

RBM models constitute an interesting class of undirected graphical
models that are thought to be useful for supervised learning tasks.
However, when viewed as generative statistical models, RBMs can be
susceptible to forms of model impropriety such as near-degeneracy,
S-instability, and uninterpretability. These model instability problems
relate to how useful the model may be for representing realistic data
generation mechanisms. In this paper, we have provided a theoretical
framework for discussing the model properties of RBMs, as well as
provided empirical evidence that these model problems can arise away
from the origin of the parameter space.

Additionally, these models are difficult to fit using a rigorous
methodology that can incorporate estimation uncertainty, due to the
dimension of the parameter space coupled with the size of the latent
variable space. We have presented three fully Bayes-principled
MCMC-based methods for fitting RBMs, which provide posterior
distributions for quantifying uncertainty in estimating model parameters
and probabilities (e.g., Figure \ref{fig:fitting-plot}) while aiming to
avoid areas of the parameter space that will potentially lead to model
impropriety. We have shown small examples where this fitting methodology
does work, but becomes quickly intractable in large data cases.

The current common practice is to use a kind of MCMC to overcome fitting
complexities, and due to the extreme flexibility in this model class,
rigorous likelihood-based fitting for a RBM can typically seek to merely
merely return the (discrete) empirical distribution for visibles.
Current fitting methodology for RBMs does not provide uncertainty
quantification for the fitted parameters, and it is unclear that they
seeks to address the issues in model impropriety detailed in this paper.
Practitioners should be aware of this and employ some form of
regularization or penalization in the fitting.
